---
title: "Unraveling the Mystery of Rarotongan Lagoon"
subtitle: "Problem Set 1"
author: "Ian Brunjes, Travis Christy, Karla Garibay"
date: "1/14/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
library(tidyverse)
library(ggpubr)
library(here)
library(patchwork)
```

# Identifying Patterns in the Data

```{r, echo=FALSE, message=FALSE}
adult_data <- read_csv(here("data", "mean_density_adult_organisms.csv"))
adult_data$year <- as.factor(adult_data$year)
young_data <- read_csv(here("data", "mean_density_young_fish.csv"))
young_data$year <- as.factor(young_data$year)
young_data <- young_data %>% filter(n_per_1_m2 != 0)
```


## 1. Patterns in Adult Organism Density across Year/Location

```{r, echo=FALSE}
organism_plots_adult_fish <- list()
adult_data_fish <- adult_data %>% filter(type == "adult fish")
for (spp in unique(adult_data_fish$species)) {
  # Filter to organism
  species_data <- adult_data_fish %>% filter(species == spp)
  
  ## Create plot for each organism
  spp_plot <- ggplot(species_data, aes(fill = year, x = island, y = n_per_100_m2),) +
    geom_bar(position = "dodge", stat = "identity") +
    geom_errorbar(
      aes(ymin = n_per_100_m2 - ci_95, ymax = n_per_100_m2 + ci_95),
      width=0.4,
      colour="orange",
      alpha=0.9,
      size=0.9,
      position = position_dodge(0.9)
    ) +
    labs(
      x = "",
      y = "",
      title = spp
    ) +
    theme(plot.title = element_text(size=10))
  
  organism_plots_adult_fish[[spp]] <- spp_plot
}
```

```{r, echo=FALSE}
wrap_plots(organism_plots_adult_fish, ncol = 2) +
  plot_annotation(title =  expression(paste("Adult Fish: Number per 100", m^{2}," of reef"))) +
  plot_layout(guides = "collect")
```


```{r, echo=FALSE}
organism_plots_adult_other <- list()
adult_data_other <- adult_data %>% filter(type != "adult fish")
for (spp in unique(adult_data_other$species)) {
  # Filter to organism
  species_data <- adult_data_other %>% filter(species == spp)
  
  ## Create plot for each organism
  spp_plot <- ggplot(species_data, aes(fill = year, x = island, y = n_per_100_m2),) +
    geom_bar(position = "dodge", stat = "identity") +
    geom_errorbar(
      aes(ymin = n_per_100_m2 - ci_95, ymax = n_per_100_m2 + ci_95),
      width=0.4,
      colour="orange",
      alpha=0.9,
      size=0.9,
      position = position_dodge(0.9)
    ) +
    labs(
      x = "",
      y = "",
      title = spp
    ) +
    theme(plot.title = element_text(size=10))
  
  organism_plots_adult_other[[spp]] <- spp_plot
}
```

```{r, echo=FALSE}
wrap_plots(organism_plots_adult_other) +
  plot_annotation(title =  expression(paste("Other Organisms: Amount of Bottom Covered (", m^{2}, ") per 100", m^{2}," of Reef"))) +
  plot_layout(guides = "collect")
```

## 2. Patterns in New Recruit Density across Year/Location

```{r, echo=FALSE}
organism_plots_young <- list()
for (spp in unique(young_data$fish)) {
  # Filter to organism
  species_data <- young_data %>% filter(fish == spp)
  habitat <- unique(species_data["microhabitat"])[[1]]
  ## Create plot for each organism
  spp_plot <- ggplot(species_data, aes(fill = year, x = island, y = n_per_1_m2),) +
    geom_bar(position = "dodge", stat = "identity") +
    geom_errorbar(
      aes(ymin = n_per_1_m2 - ci_95, ymax = n_per_1_m2 + ci_95),
      width=0.4,
      colour="orange",
      alpha=0.9,
      size=0.9,
      position = position_dodge(0.9)
    ) +
    labs(
      x = "",
      y = "",
      title = paste(spp, " (", habitat, ")", sep = "")
    ) +
    theme(plot.title = element_text(size=10))
  
  organism_plots_young[[spp]] <- spp_plot
}
```


```{r, echo=FALSE}
wrap_plots(organism_plots_young, ncol = 2, height = 3) +
  plot_annotation(title = expression(paste("Young Fish: Number of new recruits per 1", m^{2}," of preferred microhabitat"))) +
  plot_layout(guides = "collect")
```




## 3. Comparing Density of Adults and Recruits

```{r, echo=FALSE}
# Joining adult + recruit + preferred microhabitat data
adult_recruit_join <- left_join(
  adult_data_fish,
  young_data,
  by = c("species" = "fish", "year" = "year", "island" = "island"),
  suffix = c("_adult", "_recruit")
)

full_join <- left_join(
  adult_recruit_join,
  adult_data_other,
  by = c("microhabitat" = "species", "year" = "year", "island" = "island"),
  suffix = c("_species", "_habitat")
)

# Calculating values for 3. & 5.
full_join <- full_join %>% mutate(
  recruits_per_100_m2 = n_per_1_m2 * n_per_100_m2_habitat,
  adults_per_m2_microhabitat = n_per_100_m2_species / n_per_100_m2_habitat
)

```

```{r, echo=FALSE}
density_comparison_plots <- list()
for (spp in unique(full_join$species)) {
  # Filter to organism
  species_data <- full_join %>% filter(species == spp)
  
  ## Create plot for each organism
  spp_plot <- ggplot(species_data, aes(y = n_per_100_m2_species, x = recruits_per_100_m2)) +
    geom_point(aes(color = year, shape = island)) +
    labs(
      y = "number of adults",
      x = "number of recruits",
      title = spp
    ) +
    geom_smooth(method = "lm", se=FALSE, color = "black", size = 0.2) +
    stat_regline_equation(aes(label = ..rr.label..))
  
  density_comparison_plots[[spp]] <- spp_plot
}
```

```{r, echo=FALSE}
wrap_plots(density_comparison_plots) +
  plot_annotation(title = "Density of Adults versus New Recruits per 100m2 of Reef") +
  plot_layout(guides = "collect")
```

## 4. Comparing Density of Adults with Area of Preferred Microhabitat
```{r, echo=FALSE}
density_microhabitat_plots <- list()
for (spp in unique(full_join$species)) {
  # Filter to organism
  species_data <- full_join %>% filter(species == spp)
  
  ## Create plot for each organism
  spp_plot <- ggplot(species_data, aes(x = n_per_100_m2_habitat, y = n_per_100_m2_species)) +
    geom_point(aes(color = year, shape = island)) +
    labs(
      x = "area of preferred microhabitat",
      y = "number of adults",
      title = spp
    ) +
    geom_smooth(method = "lm", se=FALSE, color = "black", size = 0.2) +
    stat_regline_equation(aes(label = ..rr.label..))
  
  density_microhabitat_plots[[spp]] <- spp_plot
}
```

```{r, echo=FALSE}
wrap_plots(density_microhabitat_plots) +
  plot_annotation(title = "Area of Microhabitat versus Density of Adults per 100 m2 of Reef") +
  plot_layout(guides = "collect")
```

## 5. Density of Adult Fish per Area of Microhabitat preferred by Young

```{r, echo=FALSE}
density_per_microhabitat_plots <- list()
for (spp in unique(full_join$species)) {
  # Filter to organism
  species_data <- full_join %>% filter(species == spp)
  habitat <- unique(species_data["microhabitat"])[[1]]
  ## Create plot for each species
  spp_plot <- ggplot(species_data, aes(fill = year, x = island, y = adults_per_m2_microhabitat),) +
    geom_bar(position = "dodge", stat = "identity") +
    labs(
      x = "",
      y = "",
      title = paste(spp, " (", habitat, ")", sep = "")
    ) +
    theme(plot.title = element_text(size=10))
  
  density_per_microhabitat_plots[[spp]] <- spp_plot
}
```

```{r, echo=FALSE}
wrap_plots(density_per_microhabitat_plots) +
  plot_annotation(title =  expression(paste("Number of adult fish per ", m^{2}," of microhabitat preferred by young"))) +
  plot_layout(guides = "collect")
```

## 6. Comparing Density between each Species Pair

```{r, echo=FALSE}
# Get all possible pairs of fish species
pairs <- combn(unique(adult_data_fish$species), 2)

# For each pair, create plot
pair_plots <- list()
for (i in seq(dim(pairs)[2])) {
  pair = pairs[,i]
  s1 <- pair[1]
  s2 <- pair[2]
  
  s1_data <- adult_data_fish %>% filter(species == s1)
  s2_data <- adult_data_fish %>% filter(species == s2)
  
  s1_s2_join <- left_join(s1_data, s2_data, by = c("year" = "year", "island" = "island"), suffix = c("_s1", "_s2"))
  
  # create plot
  pair_plot <- ggplot(s1_s2_join, aes(x = n_per_100_m2_s1, y = n_per_100_m2_s2)) +
    geom_point(aes(color = year, shape = island)) +
    labs(
      x = s1,
      y = s2
    ) +
    geom_smooth(method = "lm", se=FALSE, color = "black", size = 0.2) +
    stat_regline_equation(aes(label = ..rr.label..))
  
  pair_plots[[paste(s1,s2)]] <- pair_plot
}
```

```{r, echo=FALSE}
wrap_plots(pair_plots) +
  plot_annotation(title =  "Comparing Adult Fish Densities between Species") +
  plot_layout(guides = "collect")
```



